---
title: "YOUR TITLE HERE"
subtitle: "Stat231: Google Calendar Report"
author: "Thai Nguyen"
date: "Due Friday, September 25 by 5:00 PM EST"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# How do I spend my time?

```{r}
# include your import of the data and preliminary wrangling here

# importing data
library(tidyverse)
library(lubridate)
library(ical)
path <- "D:/Study/STAT 231 temp/Git/Stat-231-repo/Calendar assignment"
filename <- "Test schedule.ics"

my_calendar0 <- ical_parse_df(file = paste0(path,"/",filename)) %>%
  mutate(start_datetime = with_tz(start, tzone = "America/New_York")
    , end_datetime = with_tz(end, tzone = "America/New_York")
    , length_hour = end_datetime - start_datetime
    , date = floor_date(start_datetime, unit = "day"))
```

```{r}
# initial data wrangling
names(my_calendar0)
my_calendar <- my_calendar0 %>%
  filter(date > "2020-09-01") %>%
  select(-c(uid, description, last.modified, status)) %>%
  rename(length_minute = length_hour) %>%
  mutate(length_hour = length_minute / 3600)

```


Describe your question(s) here.  Briefly describe your data collection process, including how you defined variables of interest.

Describe what information is conveyed through data visualization #1 (that you'll create below) here.

```{r}
# Write your code to create data visualization #1 here.  
# Be sure to label your axes and include a title to give your data's context.
```

Describe what information is conveyed through data visualization #2 (that you'll create below) here.

```{r}
# Write your code to create data visualization #2 here.  
# Be sure to label your axes and include a title to give your data's context.
```

Describe what information is conveyed through the table (that you'll create below) here.

```{r, results='asis'}
# write your code to create the table here
# if you want to make your table's appearance nicer, check out:
# the xtable package (https://cran.r-project.org/web/packages/xtable/vignettes/xtableGallery.pdf); or
# the kable function in the knitr package (https://bookdown.org/yihui/rmarkdown-cookbook/kable.html)
```

To conclude, briefly summarize what you found in response to the questions posed here.

\newpage
# Reflection

Write your one-page reflection here in paragraph form.  In particular, address:

- What difficulties in the data collection and analysis process did you encounter?  Identify two of your main hurdles in gathering accurate data.
- What implications does that have for future data collection and/or analysis projects? 
- How much data do you think you'd need to collect in order to answer your question(s) of interest? Would it be hard to collect that data? Why or why not?
- As someone who provides data, what expectations do you have when you give your data (e.g. to Facebook, Google, MapMyRun, etc.)?  
- As someone who analyzes others' data, what ethical responsibilities do you have?



