---
title: 'STAT 231: Problem Set 1A'
author: "THAI NGUYEN"
date: "due by 5 PM on Monday, August 31"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

In order to most effectively digest the textbook chapter readings -- and the  new R commands each presents -- series A homework assignments are designed to encourage you to read the textbook chapters actively and in line with the textbook's Pro Tip on page 33:

"\textbf{Pro Tip}: If you want to learn how to use a particular command, we highly recommend running the example code on your own" 

A more thorough reading and light practice of the textbook chapter prior to class allows us to dive quicker and deeper into the topics and commands during class.  Furthermore, learning a programming lanugage is like learning any other language -- practice, practice, practice is the key to fluency.  By having two assignments each week, I hope to encourage practice throughout the week.  A little coding each day will take you a long way!

*Series A assignments are intended to be completed individually.*  While most of our work in this class will be collaborative, it is important each individual completes the active readings.  The problems should be straightforward based on the textbook readings, but if you have any questions, feel free to ask me!

Steps to proceed:

\begin{enumerate}
\item In RStudio, go to File > Open Project, navigate to the folder with the course-content repo, select the course-content project (course-content.Rproj), and click "Open" 
\item Pull the course-content repo (e.g. using the blue-ish down arrow in the Git tab in upper right window)
\item Copy ps1A.Rmd from the course repo to your repo (see page 6 of the GitHub Classroom Guide for Stat231 if needed)
\item Close the course-content repo project in RStudio
\item Open YOUR repo project in RStudio
\item In the ps1A.Rmd file in YOUR repo, replace "YOUR NAME HERE" with your name
\item Add in your responses, committing and pushing to YOUR repo in appropriate places along the way
\item Run "Knit PDF" 
\item Upload the pdf to Gradescope.  Don't forget to select which of your pages are associated with each problem.  \textit{You will not get credit for work on unassigned pages (e.g., if you only selected the first page but your solution spans two pages, you would lose points for any part on the second page that the grader can't see).} 
\end{enumerate}

```{r, setup, include=FALSE}
# Load packages here 
library(mdsr)

knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small",   # slightly smaller font for code
  message=FALSE)   
```

\newpage
# 1. GDP and education

### a. 
Figure 3.3 in Section 3.1.1 shows a scatterplot that uses both location and label as aesthetics.  Reproduce this figure.  Hint: you'll need to define 'g' based on code from earlier in Section 3.1.1.

```{r}
data(CIACountries)

# define the plot object
g <- ggplot(data = CIACountries, aes(y = gdp, x = educ))
  
# print the plot 
g + geom_text(aes(label = country, color = net_users), size = 3)
```

### b.
Now, update the plot with more informative labels.  Label the x-axis "% of GDP spent on education" and the y-axis "Gross Domestic Product (GDP)".  Hint: see Section 3.2.2 for an example of one way to label the axes.

```{r}
g + geom_text(aes(label = country, color = net_users), size = 3) + ylab("Gross Domestic Product (GDP)") + xlab("% of GDP spent on education")

```

### c.
Next, move the legend so that it's located on the top of the plot as opposed to the right of the plot.  Hint: see Section 3.1.4 for an example on how to change the legend position.

```{r}
g + geom_text(aes(label = country, color = net_users), size = 3) + ylab("Gross Domestic Product (GDP)") + xlab("% of GDP spent on education") + theme(legend.position = "top")
```

### d.
Lastly, Section 3.1.2 discusses *scale*, and demonstrates how to display GDP on a logarithmic scale to better discern differences in GDP.  Update the figure so GDP is on a log10 scale.  

```{r}
g + geom_text(aes(label = country, color = net_users), size = 3) + ylab("Gross Domestic Product (GDP)") + xlab("% of GDP spent on education") + theme(legend.position = "top") + coord_trans(y = "log10")
```


\newpage 
# 2. Medical procedures

### a.
Consider Figure 3.7 in Section 3.2.1.  What does `reorder(drg, mean_charge)` do?  Recreate the plot, but use `x = drg` instead of `x = reorder(drg, mean_charge)`.  What happens?

> ANSWER: 'reorder(drg, mean_charge)' rearranges the data of Medical Procedure in order from lowest to highest Statewide Average Charges. When we use 'x = drg' instead, the data is ordered from the lowest to highest number representing the Medical Procedure instead.

```{r}
data(MedicareCharges)
ChargesNJ <- MedicareCharges %>% 
  ungroup() %>%
  filter(stateProvider == "NJ")

# create the plot object
p <- ggplot(data = ChargesNJ, aes(x = drg, y = mean_charge)) + geom_bar(fill = "gray", stat = "identity") + ylab("Statewide Average Charges ($)") + xlab("Medical Procedure (DRG)") + theme(axis.text.x = element_text(angle = 90, hjust = 1))
  
# print the plot
p 
```

### b.
Replace `x = drg` with `x = reorder(drg, mean_charge)`, but also remove the `theme()` line.  Now what happens?  What was the purpose of the `theme()` line?  Hint: You may need to knit the document and look at the pdf to better observe what's happening.

> ANSWER: The theme line rotates the number represent the Medical Procedure vertically, which makes looking at the graph easier. Without the theme() line, there are overlaps between the labeling of each column.

```{r}
g <- ggplot(data = ChargesNJ, aes(x = reorder(drg, mean_charge), y = mean_charge)) + geom_bar(fill = "gray", stat = "identity") + ylab("Statewide Average Charges ($)") + xlab("Medical Procedure (DRG)") 

g
```


\newpage 
# 3. Historical baby names

As you read through (and, better yet -- code along with (not required, but useful practice!)) -- the extended example on historical baby names in section 3.3.1, write down two questions you have about any of the R code used in that example.  (Your questions could be about what a specific part of the code -- ggplot or not -- is actually doing, or a more general question about any of the commands used.)  Please be thoughtful about your questions; we will use them (anonymously) in an exercise in class this week.

> ANSWER: I tried to code along with the text but couldn't find the Hmisc package that is required to get the median year of birth data so I skipped that part. Thus the end result is missing the blue bar and the black line. 
I was wondering why in the text, the line "The median living Joseph is 37 years old" is fully out of the arrow, while in my graph, the arrow cut into the words?

```{r}
# to get you started following along . . .
library(babynames)
BabynamesDist <- make_babynames_dist()

joseph <- BabynamesDist %>%
  filter(name == "Joseph" & sex == "M")

name_plot <- ggplot(data = joseph, aes(x = year)) +
  geom_bar(stat = "identity", aes(y = count_thousands*alive_prob)
           , fill = "#b2d7e9", color = "white")

name_plot <- name_plot + ylab("Number of People (thousands)") + xlab(NULL) + ggtitle("Age Distribution of American Boys Named Joseph") + geom_text(x = 1935, y = 40, label = "Number of Josephs\nborn each year") + geom_text(x = 1915, y = 13, label = "Number of Josephs\nborn each year\nestimated to be alive\non 1/1/2014", colour = "#b2d7e9") + 
  geom_text(x = 2003, y = 40, 
            label = "The median\nliving Josephs\nis 37 years old",
            colour = "darkgray") +
  geom_curve(x = 1995, xend = 1974, y = 40, yend = 24, 
             arrow = arrow(length = unit(0.3, "cm")), curvature = 0.5) + ylim(0, 42)

name_plot



```




