---
title: "STAT 231: Problem Set 7B"
author: "Thai Nguyen"
date: "due by 5 PM on Friday, October 30"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

This homework assignment is designed to help you futher ingest, practice, and expand upon the material covered in class over the past week(s).  You are encouraged to work with other students, but all code and text must be written by you, and you must indicate below who you discussed the assignment with (if anyone).  

Steps to proceed:

\begin{enumerate}
\item In RStudio, go to File > Open Project, navigate to the folder with the course-content repo, select the course-content project (course-content.Rproj), and click "Open" 
\item Pull the course-content repo (e.g. using the blue-ish down arrow in the Git tab in upper right window)
\item Copy ps7B.Rmd from the course repo to your repo (see page 6 of the GitHub Classroom Guide for Stat231 if needed)
\item Close the course-content repo project in RStudio
\item Open YOUR repo project in RStudio
\item In the ps7B.Rmd file in YOUR repo, replace "YOUR NAME HERE" with your name
\item Add in your responses, committing and pushing to YOUR repo in appropriate places along the way
\item Run "Knit PDF" 
\item Upload the pdf to Gradescope.  Don't forget to select which of your pages are associated with each problem.  \textit{You will not get credit for work on unassigned pages (e.g., if you only selected the first page but your solution spans two pages, you would lose points for any part on the second page that the grader can't see).} 
\end{enumerate}

```{r, setup, include=FALSE}
library(tidyverse)
library(ggnetwork)
library(igraph)

knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
```


\newpage 
# If you discussed this assignment with any of your peers, please list who here:

> ANSWER: Leah Johnson

\newpage
# 1. More Migration

1a. Consider migration between the following countries: Brazil, Ghana, Great Britain, Honduras, India, South Korea, United States, and Vietnam.  Compare the TOTAL (males + females) migration between these countries over time.  In separate (directed) graphs for 1980 and 2000, visualize the network for the these countries with edge width and/or edge color corresponding to migration flow size.  Interpret the two graphs -- what *information in context* do they convey?

> ANSWER:  It seems like migration to and from Great Britain and the US are among the largest in size in both 1980 and 2000, whereby migration to and from countries in Asia like Vietnam or South Korea is virtually non-existent unless moving to and from the two aforementioned countries. 
> From the 1980 graph, we see that the United States is the most popular destination country in 1980, with most people arriving there coming from Great Britain (more than 600,000), as well as from Korea, Vietnam, and India. The most popular path seems is between Great Britain to the USA. We can also see a large migration flow from India to Great Britain (maybe due to post-colonial migration), and a moderate migration flow from the United States to Great Britain. 
> From the 2000 graph,in general migration flows has become larger (meaning more people are migrating in comparison to 1980), with the largest flow is still from Great Britain to the US. There are fewer relatively large migration in comparison to 1980, besides the aforementioned flow and the one from India to Great Britain.


```{r, message = FALSE}
path_in <- "/Users/thainguyen/Desktop/STUDYYYYYYYYYYYYYYYYYYYYYY/College/Fall 2020/Data Science STAT 231/Git/Stat-231-repo/Data"
MigrationFlows <- read_csv(paste0(path_in,"/MigrationFlows.csv"))

countries <- c("BRA","GBR","GHA","HND","IND","KOR", "USA", "VNM")

# need migration overall: 
migration <- MigrationFlows %>%
  filter(destcode %in% countries & origincode %in% countries) %>%
  select(c(sex, destcode, origincode, Y2000, Y1980)) %>%
  # do some prelim data wrangling to combine numbers for males + females
  pivot_wider(names_from = sex, values_from = c(Y2000, Y1980)) %>%
  mutate(Y2000 = Y2000_Male + Y2000_Female,
         Y1980 = Y1980_Male + Y1980_Female) %>%
  select(c(destcode, origincode, Y2000, Y1980))

#Compare total migration
migration %>%
  arrange(desc(Y2000))
migration %>%
  arrange(desc(Y1980))


# Prepare data set for 1980 network graph
migration_1980 <- migration %>%
  # include only positive values
  filter(Y1980 > 0) %>%
  select(origincode, destcode, Y1980)

mig_1980 <- graph_from_data_frame(migration_1980, directed = TRUE)
mig_1980_network <- ggnetwork(mig_1980)

# Create 1980 graph
ggplot(data = mig_1980_network, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(aes(color = Y1980), arrow = arrow(type = "closed", length = unit(6,"pt")), 
             curvature = 0.1, size = 0.8) +
  geom_nodes() +
  geom_nodelabel_repel(aes(label = name)) +
  theme_blank() +
  labs(title = "Migration Flow for Selected Countries in 1980", 
       color = "1980 Migration Flow") +
  scale_color_continuous(type = "viridis")

# prepare data set for 2000 network graph
migration_2000 <- migration %>%
  # include only positive values
  filter(Y2000 > 0) %>%
  select(origincode, destcode, Y2000)

mig_2000 <- graph_from_data_frame(migration_2000, directed = TRUE)
mig_2000_network <- ggnetwork(mig_2000)

#Create 2000 graph
ggplot(data = mig_2000_network, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(aes(color = Y2000), arrow = arrow(type = "open", length = unit(6,"pt")), 
             curvature = 0.1, size = 0.7) +
  geom_nodes() +
  geom_nodelabel_repel(aes(label = name)) +
  theme_blank() +
  labs(title = "Migration Flow for Selected Countries in 2000", 
       color = "2000 Migration Flow") +
  scale_color_continuous(type = "viridis")
```

1b. Compute the *unweighted* in-degree for Brazil in this network from 2000, and the *weighted* in-degree for Brazil in this network from 2000.  In 1-2 sentences, interpret these numbers in context (i.e., without using the terms "in-degree" or "weighted").

> ANSWER:  The unweighted in-degree for Brazil in 2000 is 7, which indicates that there are 7 edges directed towards the node corresponding to Brazil. This means that there was migration flow from 7 of the other countries to Brazil in 2000. The weighted in-degree for Brazil in 2000 is 20885. This takes into account the weight of each edge directed towards Brazil as well, which means that the number of people migrating to Brazil from those 7 edges above total up to 20885, with each edge represent a different number of migration flow. 

```{r}
# unweighted 
degree(mig_2000, v = "BRA", mode = "in")

# weighted 
strength(mig_2000, v = "BRA", weights = E(mig_2000)$Y2000, mode = "in")
```


1c. Among these same countries, idenitify the top 5 countries *of origin* and *of destination* (separately) in 1980 using (weighted) degree centrality.  Interpret this information.

> ANSWER: Using the weighted degree centrality, which takes into account the weight of each edge, which represents the number of people migrating, we see that Great Britain is the top country of origin in 1980 amongst the countries above (which is understandable since the yellow edge was directed outward from Great Britain in our network). It is followed by India, Korea, and Vietnam (which we also mentioned in our network analysis). The United States was also a popular country of origin, albeit only at the 5th position. On the other hand, the United States is actually the top destination country in 1980 (noted earlier), when looking at the network as the largest weighted edges are all pointing towards it. Great Britain is in second place,as well as Brazil, India, and Korea, which are interesting as we could not identify as easily from our network visualization above.


```{r}
V(mig_1980)$out_degree <- strength(mig_1980, weights = E(mig_1980)$Y1980, 
                                  mode = "out") 
V(mig_1980)$in_degree <- strength(mig_1980, weights = E(mig_1980)$Y1980, 
                                  mode = "in") 

degree_df_1980 <- cbind.data.frame(name = V(mig_1980)$name, 
                 out_degree = V(mig_1980)$out_degree, 
                 in_degree = V(mig_1980)$in_degree)

# Top 5 of origin in 1980
degree_df_1980 %>%
  select(name, out_degree) %>%
  arrange(desc(out_degree)) %>%
  head(5)

# Top 5 of destination in 1980
degree_df_1980 %>%
  select(name, in_degree) %>%
  arrange(desc(in_degree)) %>%
  head(5)
```


1d. Among these same countries, idenitify the top 5 countries *of origin* and *of destination* (separately) in 2000 using (weighted) degree centrality.  Interpret this information.

> ANSWER: We see that Great Britain is the top country of origin in 2000 as well (which is understandable since the yellow edge was directed outward from Great Britain in our network in 2000 as well). It is followed by India, and something we should take note of is the outward migration drops from 630 thousand in 1980 to only 206 thousand in 2000, only 1/3 of what it was before. The United States is the third most popular country of origin in 2000, which is an improvement in terms of ranking as it was fifth in 1980, and it is followed by Brazil and Vietnam. (The US got this position not because more people migrating outward, but rather people in other countries migrate outward less). Also, Korea is no longer in the top 5. 
> On the other hand, the United States is the top destination country in 2000 as well. So the most people seemed to go to the United States, even though the number of those migrating drop by almost half, from 1.7 million to only 934 thousand. This is followed by Great Britain, Brazil, India, and surprisingly Ghana, which did not show up in any of the top 5 positions (either in origin or destination) in 1980. 


```{r}
V(mig_2000)$out_degree <- strength(mig_2000, weights = E(mig_2000)$Y2000, 
                                  mode = "out") 
V(mig_2000)$in_degree <- strength(mig_2000, weights = E(mig_2000)$Y2000, 
                                  mode = "in") 

degree_df_2000 <- data_frame(name = V(mig_2000)$name, 
                 out_degree = V(mig_2000)$out_degree, 
                 in_degree = V(mig_2000)$in_degree)

# Top 5 of origin in 2000
degree_df_2000 %>%
  select(name, out_degree) %>%
  arrange(desc(out_degree)) %>%
  head(5)

# Top 5 of destination in 2000
degree_df_2000 %>%
  select(name, in_degree) %>%
  arrange(desc(in_degree)) %>%
  head(5)
```


1e. What is the diameter of this network in 2000? In 1-2 sentences, interpret this value.

> ANSWER:  The diameter of the network in 2000 is 2. This means that the maximum number of paths between any two nodes is 2, meaning that any two countries are separated by at most one other country in terms of migration flow between them. 


```{r}
diameter(mig_2000, directed = TRUE)
```


1f. What is the density of this network in 2000? In 1-2 sentences, interpret this value.

> ANSWER:  The density of the network tells us the proportion of all possible edges that are actually used. The density of this network in 2000 is 0.768, which indicates that of all of the possible migration paths between the countries, 76.8\% of them were actually used in the year 2000.

```{r}
# note that number of possible edges = n(n-1), where n is number of nodes
ecount(mig_2000)/(vcount(mig_2000)*(vcount(mig_2000) - 1))
```


\newpage
# 2. Love Actually (OPTIONAL PRACTICE)

This problem is *optional* and will not be graded, but is given to provide additional practice interpreting networks and as another real-world example of network analysis that might be intriguing to film buffs. 

Consider the figure "The Two Londons of 'Love Actually'" in [this FiveThirtyEight article](https://fivethirtyeight.com/features/the-definitive-analysis-of-love-actually-the-greatest-christmas-movie-of-our-time/). 

2a. Based on this figure, is the network connected?  In 1-2 sentences, please explain.

> ANSWER: 

2b. Based on the figure, what is the (unweighted) degree for Emma Thompson?  What is the (unweighted) degree for Keira Knightley?  Explain what these values mean for these characters.

> ANSWER: 

2c. Based on the figure, for whom would the (unweighted) betweenness centrality measure be higher: Colin Firth or Hugh Grant?  Explain what this implies. 

> ANSWER:  